---
title: Syllabus
slug: syllabus
citeproc: true
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    toc: true
    fig_caption: false
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tibble)
library(pander)
```

`r blogdown::shortcode("courseinfo")`

## Course content 

Bayesian analysis is a statistical tool that is becoming increasingly popular in medical science. Notably, Bayesian approaches have become commonly used in adaptive designs for Phase I/II trials, in meta- analyses, and also in transcriptomics analysis. This course provides an introduction to Bayesian tools, with an emphasis on biostatistics applications, in order to familiarize students with such methods and their practical applications. Thanks to its rich modeling possibilities the Bayesian framework is appealing, especially when the number of observations is scarce. It can adaptively incorporate information as it becomes available, an important feature for early phase clinical trials. Thus, adaptive Bayesian designs for Phase I/II trials reduce the chances of unnecessarily exposing participants to inappropriate doses and have better decision-making properties compared to the standard rule-based dose-escalation designs. Besides, the use of a Bayesian approach is also very appealing in meta-analyses because of: i) the often relatively small number of studies available, ii) its flexibility, iii) and its better handling of heterogeneity from aggregated results, especially in network meta-analyses. Thanks to modern computing tools, practical Bayesian analysis has become relatively straightforward, which is contributing to its increasing popularity. JAGS is a flexible software interfaced with R, that allows to easily specify a Bayesian model and that automatically perform inference for posterior parameters distributions as well as graphic outputs to monitor the quality of the analysis.

The aim of the course is to provide insights into Bayesian statistics in the context of medical studies. We will cover the following topics: 
 
 1. Bayesian modeling (prior, posterior, likelihood, Bayes theorem)
 2. Bayesian estimation (Credibility Intervals, Maximum a Posteriori, Bayes factor)
 3. Adaptive designs for Phase I/II trials
 4. Bayesian meta-analyses
 5. Practical Bayesian Analysis with `r icon::fa_r_project(colour = "#0085d6")` and `JAGS` softwares
 6. critical reading of medical publications. 
 
 All concepts will be illustrated with real-life examples from the medical literature



## Course objectives

A student who has met the objectives of the course will be able to:

 - Familiarize oneself with the **Bayesian framework**:
    1. understand and assess a Bayesian modeling strategy, and discuss its underlying assumptions 
    2. rigorously describe expert knowledge by a quantitative prior distribution 

 - Study and perform Bayesian analyses in **biomedical applications**:
    1. understand, discuss and reproduce a Bayesian (re-)estimation of a Relative Risk
    2. perform a Bayesian regression using [`r icon::fa_r_project(colour = "#0085d6")`](https://cran.r-project.org/), applied to meta-analysis
    3. understand and explain an adaptive design for Phase I/II trials and the associated decision-rule 


## Targeted audience

This course is targeted towards medical students in graduate programms at the Faculty of Health and Medical Sciences. To be able to follow this course, participants need both basic knowledge in statistics (most notably some familiarity with Maximum Likelihood Estimation) and basic knowledge in R (functional programming, vector allocation, linear regression).

Advanced mathematical training is not required as we will explain the methods on an elementary mathematical level. During the practicals on their laptop, the students will learn how to technically apply the Bayesian tools on real data, and should be able to perform a Bayesian regression by the end of the course. Note that several statistical software can be used for Bayesian analysis, however solutions will be provided for the statistical softwares `r icon::fa_r_project(colour = "#0085d6")` and `JAGS` only (alternatives such as [`WinBUGS`](https://www.mrc-bsu.cam.ac.uk/software/bugs/the-bugs-project-winbugs/) or [`STAN`](https://mc-stan.org/) will not be covered).

## Technical tools

### R and RStudio

The practicals in this class will use R extensively ! Please make sure you have an up-to-date working installation of R:

 - latest version of `r icon::fa_r_project(colour = "#0085d6")` (≥ 4.0.0) `r emo::ji("backhand index pointing right")` https://cran.r-project.org/ (watch [here](https://vimeo.com/203516510) how to install it)
 - latest version of RStudio (≥ 1.3.959) `r emo::ji("backhand index pointing right")` https://www.rstudio.com/products/rstudio/download/#download

If you are not fluent in R yet, you can have a look at those various (free) ressources on R beforehand:

 - [Teacups Giraffe turorial](https://tinystats.github.io/teacups-giraffes-and-statistics/01_introToR.html) intro class by Hasse Walum & Desirée De Leon
 - [Get Your R Act Together](https://stat545.com/r-basics.html) intro by Jenny Brian
 - [R for data science](https://r4ds.had.co.nz/) book by Hadley Whickam
 - Lessons 1 to 9 of the *R Programming* course from [SWIRL](https://swirlstats.com/students.html)
```{r echo=FALSE, eval=FALSE}
install.packages("swirl")
library(swirl)
install_course("R Programming")
swirl()
```

 - more learning ressources here: https://education.rstudio.com/learn/beginner/

### `JAGS` software

 - install the `JAGS` software from here `r emo::ji("backhand index pointing right")` https://sourceforge.net/projects/mcmc-jags/files/
 - install the `rjags` package in `r icon::fa_r_project(colour = "#0085d6")` by running the following command:
    ```{r echo=TRUE, eval=FALSE}
    install.packages("rjags")
    ```
    make sure it works by running the following command:
    ```{r echo=TRUE, eval=FALSE}
    library(rjags)
    ```
    If it *does* work correctly, you should get the following output:
    ```{r echo=FALSE, eval=TRUE, message=FALSE, results='hide'}
    Sys.setenv(LANG = "en_US.UTF-8")
    ```
    ```{r echo=FALSE, eval=TRUE, collapse=TRUE}
    library(rjags)
    ```
    ```{r echo=FALSE, eval=TRUE, message=FALSE, results='hide'}
    Sys.setenv(LANG = "fr_FR.UTF-8")
    ```

## Class chatroom

I have setup a class chatroom on [Slack](https://introbayesmed20.slack.com). This workspace is intended so that we can have asynchronous communication, but also so that you (students) can be in contact and communicate with one another.


```{r, eval=FALSE, echo=FALSE}
library(showtext)
font_add_google("Fira Sans Condensed", "fira")
showtext_auto()
p <- sticker("~/Downloads/Thomas_Bayes.jpg", package="Intro Bayes\nMed", 
        s_x=1, s_y=1, s_width=0.65,
        p_color = "#006aab", p_size=8, p_y=0.6,
        p_family = "fira",
        h_color="#0085d6", h_fill="grey95",
        filename="static/img/introbayesmed_hires.png")
```

